---
title: "Marmoset Proteomics"
output: 
  html_document:
    keep_md: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```
### Load Packges
```{r chunk1, warning=FALSE, message=FALSE, error=FALSE}

library(clusterProfiler)
library(lattice)
library(janitor)
library(skimr)
library(scales)
library(glue)
library(broom)

library(ggprism)
library(ggsci)
library(ggpubr)
library(export)

library(data.table)
library(tidyverse)
library(plotly)


theme_half_border <- theme_pubr() + 
     theme(
      strip.background = element_blank(),
      strip.text = element_text(face="bold", size = 9),
      #panel.border = element_rect(fill = NA, colour = "black"),
      axis.title.y = element_text(face="bold", size = 11),
      axis.ticks.x = element_blank(), 
      axis.line = element_line(colour = "black"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      panel.background = element_blank(),
      legend.position = "none")

```

### Read in data
```{r chunk2, warning=FALSE, message=FALSE, error=FALSE }
# Read in data exported from ScaffoldDIA
file <- "./data/log2 excl int Samples Table of M2021-074 Males search elib from 20210305.csv"
data <- fread(file, data.table=F)
head(data %>% as_tibble())

# Read in the uniprot id list that from ID mapping(https://www.uniprot.org/uploadlists/)
uniprot <- "./data/uniprot-yourlist_M20210312A94466D2655679D1FD8953E075198DA813C2215.tab"
uniprot_ann <- fread(uniprot, data.table = FALSE) %>%
   dplyr::select(-1,  uniprot_id = Entry) %>% 
   clean_names()
head(uniprot_ann %>% as.tibble())

# check how many protein does not have a gene name      
subset(uniprot_ann, gene_names=="") %>% dim() %>% .[1] #372

```

### Filter proteins (p < 0.05, fold change > 2 or < 0.5   )
```{r chunk3, warning=FALSE, message=FALSE, error=FALSE }

df_clean <- data %>% 
  clean_names() %>% 
  select(protein_name, pvalue = t_test_cl_group, old_males_rapa, fold_change = column1)  %>%
  mutate(protein_name = str_extract(protein_name, "[A-Z0-9]{2,}")) %>% 
  filter((str_detect(old_males_rapa, "^[0-9-]"))) %>% 
  mutate(old_males_rapa = as.numeric(old_males_rapa)) %>% 
  filter((str_detect(pvalue, "^[0,1]"))) %>%  
  filter(pvalue < 0.05) %>% 
  filter(old_males_rapa > 1 | old_males_rapa < -1)  %>% 
  arrange(desc(fold_change)) %>% 
  select(protein_name, fold_change)

filt_protein <- df_clean$protein_name
  
write_csv(df_clean, file = "df_clean.csv")

```


```{r chunk4,  eval=FALSE, include=FALSE}
# AnnotationHub

data(geneList, package = "DOSE")

d <- bitr(names(gene_list), 
      fromType = "UNIPROT",
      toType = "ENTREZID", 
      OrgDb = "org.Hs.eg.db"
      )


library(AnnotationHub)

hub <- AnnotationHub()
hub

unique(hub$dataprovider)
unique(hub$species) %>% grep("jacchus", ., value = TRUE)
unique(hub$rdataclass)

 dm <- query(hub, c("OrgDb", "jacchus"))

marmoset_orgb <- dm[["AH80634"]]

columns(marmoset_orgb)
length(keys(marmoset_orgb))
keytypes(marmoset_orgb)

gene_list <- data_wide %>% pull(gene_names)  #4978


entrizid_list <- bitr(gene_list, 'SYMBOL', "ENTREZID", marmoset_orgb) 
  
colnames(entrizid_list)[1] <- "gene_names"



df_id <- data_wide %>% 
  dplyr::select(uniprot_id, gene_names) %>% 
  inner_join(entrizid_list, by = "gene_names" ) 

sample_gene <- df_id$ENTREZID[1:100]


enrich.go = enrichGO(
     gene = sample_gene,
     OrgDb = marmoset_orgb,
     
     
    
    pvalueCutoff = 1,
    qvalueCutoff = 1,
    readable     = T)
     
)

barplot(enrich.go, showCategory = 20)

 enrich.go@result %>%  View
   fiter()
   mutate(GeneRatio = eval(parse(text = GeneRatio)))%>% 
   ggplot(aes(x = GeneRatio, y = fct_reorder(Description, GeneRatio))) + 
        geom_point(aes(color = p.adjust)) +
        theme_bw(base_size = 14) +
        scale_colour_gradient(limits=c(0, 0.10), low="red") +
        ylab(NULL) 


write_csv(gene_list, file = "gene_list.csv")
len

```


### Data clean
```{r chunk5}

df_clean <- fread("./df_clean.csv", data.table = FALSE) %>% as_tibble()
filt_protein <- df_clean$protein_name

all_csv <- "./data/excl int Samples Table of M2021-074 ALL search elib from 20210305.csv"

data_all <- fread(all_csv, data.table = FALSE)

head(data_all %>% as_tibble())

data_all <-
  data_all %>%
  clean_names() %>% 
  select(star, protein_name,  starts_with(c("young","old"))) %>%  
  mutate(uniprot_id = str_extract(protein_name, "[A-Z0-9]{3,}")) %>% 
  relocate(uniprot_id, .after = "protein_name") 


data_all_wide <-   data_all %>% 
  # replace "missing value as 0
  mutate(
     across(
        starts_with(c("young", "old")), 
        ~str_replace_all(., "Missing Value", "0" )
     )
   ) %>% 
 left_join(uniprot_ann, by = "uniprot_id")  %>% 
  mutate(
    across(
      starts_with(c("young", "old")),
      as.numeric
    )
  ) 


data_long <- data_all_wide %>% 
  pivot_longer(
    cols = starts_with(c("young", "old")),
    names_to = c("age", "sex", "drug", "animal_id"),
    names_sep = "_",
    values_to = "exclusive_intensity"
  ) %>% 
  unite("treatment", age:drug, sep= "_") 

slice_sample(data_long, n=8) %>% as_tibble()




data_long2 <- data_long %>% 
  mutate(log2_intensity = log2(exclusive_intensity))

# Check the distribution of exclusive intensity

histogram(~ exclusive_intensity | animal_id , data = data_long,
          xlab = "Animal ID",
          main = "Exclusive Intensity")

histogram(~ log2_intensity | animal_id , data = data_long2,
          xlab = "Animla ID",
          main = "Log2 of Exclusive Intensity")

bwplot(log2_intensity ~ animal_id , data = data_long2,
       xlab = "Animal ID",
       main = "Log2 of Exclusive Intensity")

```



### Plotting...plotly
```{r chunk6_plotly}

# functions---------------------

plotly_box <- function(data_long){
  # draw a ggplot and convert it a ggplotly object
  # 
  p <-  data_long %>% 
    ggplot(aes(treatment, 
               exclusive_intensity, 
               color = treatment,
               text = paste('<b>Protien Name:</b>', protein_names,
                            '<br><b>Gene Name: </b>', gene_names
                            
                          )       
               )
    ) +
    geom_boxplot(aes(fill = treatment), alpha = 0.1, show.legend = FALSE) +
    geom_jitter(width =  0.1, alpha = 0.3) +
    facet_wrap(~uniprot_id, scales = "free_y", ncol = 4) +
    labs(
      x = "",
      y = "Normalized Exlusive Intensity\n\n",
      title = ""
    ) + 
    scale_y_continuous(labels = scientific) +
    theme_half_border +
    theme(
      axis.text.x = element_text(angle = 45, hjust = .5, vjust = .5)
    )
  
  
  p1 <- ggplotly(p, tooltip = "text")
  return(p1)
} 
split_vector <- function(vector, chunk_length) {
  # split a vector into a sub-vector list with specific length
  # vector: a vector
  # chunk_length: the size of the sub-vector
    subgroup_list <- split(vector,
                           ceiling(seq_along(vector) / chunk_length)
                          )
    
  return(subgroup_list)
}

# split all proteins into 19 vectors(each vector contains 12 proteins)------------
subgroup_list <- split_vector(filt_protein, 12)

# plotting 19 plots----------------
plot_list_all <- lapply(subgroup_list, function(x) {
  data_long %>% 
    filter(uniprot_id %in% x) %>% 
    plotly_box()
})


# data_long %>% 
#     filter(uniprot_id %in% subgroup_list[[1]]) %>% 
#     plotly_box() 



# save 19 plottings into 19 html file---------------
walk2(plot_list_all, names(plot_list_all), function(x,y){
  html_name = paste0("All_Index_", y, ".html")
  print(html_name)
  htmlwidgets::saveWidget(x, html_name)
  }
  
) 
     

Robjects_description <- tribble(
  ~description, ~name,
  "./data/log2 excl int Samples Table of M2021-074 ALL3 search elib from 20210305.csv", "data",
  "./data/uniprot-yourlist_M20210312A94466D2655679D1FD8953E075198DA813C2215.tab", "uniprot_ann",
  "./data/excl int Samples Table of M2021-074 ALL search elib from 20210305.csv", "data_male",
  "wide_format data",                                                             "data_wide",
  "long format data",                                                        "data_long",
  "dd: uniprot_id, gene_names, ENTRIZID",                                    "df_id",               
  "list with about 200 plots",                                               "plot_list_all"
)

Robjects_description



```
### Plots for ASAH1, PCK1...
```{r chunk7, warning=FALSE, message=FALSE, error=FALSE }

id_4_plotting <- c("F7H4U3","A0A2R8M3L6", "F7ISX5", "F7I3Y0","A0A5F4WAL2", "A0A2R8MHZ1")

data_plot <- data_long %>% 
  filter(uniprot_id %in% id_4_plotting ) 

p_4presentation <- data_plot %>%    
    ggplot(aes(treatment,  exclusive_intensity,  fill = treatment)) +
    geom_boxplot(aes(fill = treatment), alpha = 0.5) +
    geom_jitter(aes(color="black", fill = treatment),width =  0.1, alpha = 0.9, show.legend = FALSE) +
    facet_wrap(~factor(gene_names, levels = c("ASAH1", "PCK1", "PLPP1", "AHSG", "CHPT1", "PKN2")),
                       scales = "free_y", ncol = 6) +
    labs(
      x = "",
      y = "Normalized Exlusive Intensity",
      title = ""
    ) + 
    scale_y_continuous(labels = scientific) +
    scale_color_aaas() + 
    scale_fill_aaas() + 
    
    theme(
      strip.background = element_blank(),
      strip.text = element_text(face="bold", size = 9),
      #panel.border = element_rect(fill = NA, colour = "black"),
      axis.title.y = element_text(face="bold", size = 11),
      axis.ticks.x = element_blank(), 
      axis.line = element_line(colour = "black"),
      #axis.text.x = element_text(angle = 45, hjust = .5, vjust = .5),
      axis.text.x = element_blank(),
      axis.text.y = element_text(face ="bold", size = 10),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      panel.background = element_blank(),
      legend.position = "bottom",
      # Change legend background color
      #legend.background = element_rect(fill = "darkgray"),
      #legend.key = element_rect(fill = "lightblue", color = NA),
      # Change legend key size and key width
      legend.key.size = unit(1, "cm"),
      legend.key.width = unit(0.5,"cm")
      
    )
p_4presentation

graph2ppt(p_4presentation, file = "marmoset_proteomics_boxplots_presentation.pptx")

data_plot %>% 
  select(gene_names, uniprot_id, treatment, animal_id, exclusive_intensity) %>% 
  unite(sample_id, treatment:animal_id, sep = "_") %>% 
  pivot_wider(
    gene_names:uniprot_id,
    names_from = sample_id,
    values_from = exclusive_intensity
  ) %>% 
  write_csv( file = "data_final_plot_wide.csv")


library(rstatix)

data_anova <- data_plot %>% 
  select(gene_names, uniprot_id, treatment, animal_id, exclusive_intensity)
  

aov_res <- data_anova %>% 
  group_by(gene_names) %>% 
  anova_test(exclusive_intensity ~ treatment)

aov_tabl <- get_anova_table(aov_res, correction = "auto")

data_anova %>% 
  group_by(gene_names) %>% 
  tukey_hsd(exclusive_intensity ~ treatment)

data_plot_split <- data_plot %>% 
  select(gene_names, uniprot_id, treatment, animal_id, exclusive_intensity) %>% 
  unite(sample_id, treatment:animal_id, sep = "_") %>% 
  pivot_wider(
    gene_names:uniprot_id,
    names_from = sample_id,
    values_from = exclusive_intensity
  ) %>% 
  group_split(gene_names)

walk2(data_plot_split, seq_along(data_plot_split), ~write_csv(.x, file = paste0("csv", .y, ".csv")))

  

```


```{r chunk8, warning=FALSE, message=FALSE, error=FALSE }

dnest <- data_long %>% group_by(uniprot_id, treatment) %>% nest()

dnest$data[[1]]


```




### PCA analysis
```{r chunk9-PCA, warning=FALSE, message=FALSE, error=FALSE }

data_pca <- data_all_wide %>% 
  select(uniprot_id, starts_with(c("young", "old"))) %>% 
  pivot_longer(
    -uniprot_id,
    names_to = "sample_marmoset",
    values_to = "exclusive_intensity"
  ) %>% 
  pivot_wider(
    id_cols = sample_marmoset,
    names_from = uniprot_id,
    values_from = exclusive_intensity
  ) 

data_pca[1:5, 1:5]


# deal with the error message-----------------------------
# 
# when run mutate(pca = map(data, ~promp))
# 
# Error: Problem with `mutate()` input `pca`.
# x 'x' must be numeric
#ℹ Input `pca` is `map(...)`.
# Run `rlang::last_error()` to see the erroe occurred
# 
# Rerun the procomp directly prcomp(data_pca, scales = T) get the following
# 
# #Error in prcomp.default(oopsmat, scale. = T) :
#cannot rescale a constant/zero column to unit variance
#
# see solution: https://blog.csdn.net/qq_39306047/article/details/112129546
# 1. check constant columns
# 2. check NAs
# 3. Is numeric data
# 4. which(apply(data, 2, var)==0)


col_exclude <- which(apply(data_pca[, -1], 2, var) == 0)
col_exclude <- names(col_exclude)

col_all <- colnames(data_pca)
col_keep2 <- col_all[!(col_all %in% col_exclude)]

col_keep <- setdiff(colnames(data_pca), col_exclude)

identical(col_keep, col_keep2)


# PCA  analysis-----------------
tidy_pca_nest <- data_pca %>% 
  select(sample_marmoset, all_of(col_keep)) %>% 
  nest(data = everything()) %>% 
  mutate(
    pca = map(data, 
              ~prcomp(.x %>% select(-sample_marmoset), center = TRUE, scale = TRUE)),
    pca_aug = map2(pca, data, ~augment(.x, data = .y))
  )

tidy_pca_nest$pca %>% 
  map(~tidy(.x, data = .y, "pcs")) %>% 
  as.data.frame() %>% 
  ggplot(aes(factor(PC), percent, group = 1)) +
  #geom_bar(stat = "identity") +
  geom_point() +
  geom_line(color = 'red') +
  labs(
    x = "Principle Components",
    y = "Variance Explained",
    title = "Scree plot: PCA"
  ) + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title = element_text(face = 'bold'))

library(ggsci)
p_pca <- 
  tidy_pca_nest %>% 
  unnest(pca_aug) %>% 
  separate(sample_marmoset, 
           into = c("age", "sex", "drug", "animal_id"), sep = "_") %>% 
  unite(group, age:drug, sep= "_") %>% 
  
  ggplot(aes(.fittedPC1, .fittedPC2)) +
  geom_point( aes(color = group, shape = group),size = 4, alpha = 0.9) + 
  labs(
    x = "PC1",
    y = "PC2",
    title = "PC score"
  ) + 
  scale_y_continuous(expand = expansion(mult = c(.1, .2)))+
  scale_fill_aaas() +
  scale_color_aaas() + 
  theme_prism(border = TRUE) +
  theme(legend.position = "bottom")

p_pca

graph2ppt(p_pca, file = "marmoset_proteomics_PCA.pptx")



```


```{r chunk10, include=FALSE, eval=FALSE}
# testing -----------------
data_long <- data_long %>% 
  filter(uniprot_id %in% filt_protein) 

data_text <- data_long %>% 
  filter(treatment == "ctrl") %>% 
  group_by(uniprot_id, pvalue, protein_names, gene_names) %>%
  summarise(
    n =n(),
    y= quantile(exclusive_intensity, 0.75)*2
  ) %>% 
  ungroup() %>% 
  mutate( treatment = "ctrl")
  


p2 <- filter(data_long, uniprot_id %in% filt_protein[1:18]) %>% 
plotly_box()


htmlwidgets::saveWidget(p2, "index.html")

```




```{r chunk11, include=FALSE, eval=FALSE}

# How to avoid warning when introducing NAs by coercion https://stackoverflow.com/questions/14984989/how-to-avoid-warning-when-introducing-nas-by-coercion
as.num = function(x, na.strings = "NA") {
    stopifnot(is.character(x))
    na = x %in% na.strings
    x[na] = 0
    x = as.numeric(x)
    x[na] = NA_real_
    x
}



set.seed(1)
vot <- rnorm(100,15,5)
vot <- sort(vot, decreasing=FALSE)
phon <- "t"
stress <- c(rep("stressed",50),rep("unstressed",50))

df1 <- data.frame(phon,stress,vot)

vot <- rnorm(100,-60,15)
vot <- sort(vot, decreasing=FALSE)
phon <- "d"
df2 <- data.frame(phon,stress,vot)

df3 <- rbind(df1, df2)
summary(df)


bwplot(vot ~ phon | stress, data= df3)

```