---
title: "downstream_analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Reference: 
  - http://evomics.org/wp-content/uploads/2016/01/phyloseq-Lab-01-Answers.html
  - https://yulab-smu.top/MicrobiotaProcessWorkshop/articles/MicrobiotaProcessWorkshop.html
## \\comp-srvr147142\Tranfer\Zhang_Research\0-Experiments_YehLab\Proj_Marmoset Study\Marmoset_mICROBIOME\Microbiome\  
  


```{r chunk1, warning=FALSE, message=FALSE, error=FALSE }

library(data.table)
library(tidyverse)
library(phyloseq)
library(ggtree)
library(ggtreeExtra)
library(MicrobiotaProcess)

```

### Prepare data
```{r chunk2, warning=FALSE, message=FALSE, error=FALSE }

if (!exists("sample_info")) {
  
  sample_info <- read_csv("./data/sample_info.csv")
  sample_df <-  sample_info %>% 
  select(sample, age, gender, group, customer_label) %>% 
  column_to_rownames(var = "sample")   
  save(sample_df, file = "sample_df.Rdata")
  
}

data_info <- tibble::tibble(
           rdata = c("taxa_tbl.Rdata", "sample_df.Rdata","seqtab.nochim.Rdata" ),
           df_name = c("taxa", "sample_df", "seqtab.nochim")
           )


map2(data_info$df_name, data_info$rdata, ~(if (!exists(.x)){load(.y, .GlobalEnv)}))

# load("taxa_tbl.Rdata")
# load("sample_df.Rdata")
# load("seqtab.nochim.Rdata")

```


### Construct a phyloseq object without using import_dada2
```{r chunk3, include=FALSE}

ps <- phyloseq(
        otu_table(seqtab.nochim, taxa_are_rows = FALSE),
        sample_data(sample_df),
        tax_table(taxa)
      )

ps

# the below procedures have been covered if use import_dada2 function)
dna <- Biostrings::DNAStringSet(taxa_names(ps))
names(dna) <- taxa_names(ps)
ps <- merge_phyloseq(ps, dna)
taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))

ps

plot_richness(ps, x = "group", measures=c("Shannon", "Simpson"), color="group")

```

### import data using import_dada2
```{r chunk4, warning=FALSE, message=FALSE, error=FALSE }


if (all(data_info$df_name %in% ls())) {
  ps_dada2 <- import_dada2(
            seqtab = seqtab.nochim,
            taxatab = taxa,
            sampleda = sample_df
  )
  
  ps_dada2
}
```


### venn diagram
```{r chunk5, warning=FALSE, message=FALSE, error=FALSE }

library(vegan)
library(coin)
library(ggnewscale)
library(VennDiagram)
library(UpSetR)

vennlist <- get_vennlist(obj=ps_dada2, factorNames="age")
upsetda <- get_upset(obj=ps_dada2, factorNames="age")

vennp <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("#00AED7", "#FD9347"),
                      cat.col=c("#00AED7", "#FD9347"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")
grid.newpage()
venn_age <- grid::grid.draw(vennp)



vennlist_gender <- get_vennlist(obj=ps_dada2, factorNames="gender")
upsetda_gender <- get_upset(obj=ps_dada2, factorNames="gender")

vennp_gender <- venn.diagram(vennlist_gender,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("#00AED7", "#FD9347"),
                      cat.col=c("#00AED7", "#FD9347"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")
grid.newpage()
grid::grid.draw(vennp_gender)

```


### Summarize sequencing depth, in general and by category
```{r chunk6, warning=FALSE, message=FALSE, error=FALSE}

sdt <- data.table(
         as(sample_data(ps_dada2), "data.frame"),
         TotalReads = sample_sums(ps_dada2),
         keep.rownames = TRUE
       )

setnames(sdt, "rn", "SampleID")

pSeqDepth <- ggplot(sdt, aes(TotalReads)) +
  geom_histogram(bins = 100) + 
  ggtitle("Sequencing Depth")

pSeqDepth + facet_grid(age ~ gender)

pSeqDepth


```


### taxa total counts histogram
```{r chunk7, warning=FALSE, message=FALSE, error=FALSE }

tdt <- data.table(
        tax_table(ps_dada2),
        TotalCounts = taxa_sums(ps_dada2),
        OTU = taxa_names(ps_dada2)
)

ggplot(tdt, aes(TotalCounts)) +
  geom_histogram() +
  ggtitle("Histogram of Total Counts")


# taxa cumulative sum

taxcumsum <- tdt[, .N, by = TotalCounts]
setkey(taxcumsum, TotalCounts)
taxcumsum[, CumSum := cumsum(N)]

p_cumsum <- ggplot(taxcumsum, aes(TotalCounts, CumSum)) +
  geom_point() +
  labs(
    x = "Filtering Threashhold, Minimal Total Counts",
    y = "OTUs Filtered",
    title = "OTUs that would be filtered vs the minimum count threshhold"
  )

p_cumsum + xlim(0,100)

``
### Data exploration and filtering


# explore data for setting criteria for filtring----
sdt <- data.frame(
  as(sample_data(ps_dada2), "data.frame"),
  total_reads = sample_sums(ps_dada2)
 )



sdt  %>% 
  ggplot (aes(total_reads, fill = gender)) +
    geom_histogram(bins = 40) +
    geom_text(aes(y = 1, label = customer_label),

              nudge_y = 0.05, nudge_x = 0,
    ) +
  
    ggtitle("Sequencing Depth") 

library(lattice)

lattice::barchart( total_reads ~ factor(customer_label) | gender * age ,
                   data= sdt, layout = c(1,4))

                   
tdt <- data.table(
    tax_table(ps_dada2), 
    total_counts = taxa_sums(ps_dada2),
    OTU = taxa_names(ps_dada2)
)



tdt[(total_counts <= 0), .N]
tdt %>% filter(total_counts < 0) %>% nrow()

taxcumsum <-  tdt[, .N, by = total_counts] %>% 
  setkey(total_counts)  %>% 
  .[, CumSum := cumsum(N)]

tdt %>% count(total_counts) %>% 
  mutate(cumsum = cumsum(n)) %>% head



pCumSum <- taxcumsum %>%
  ggplot(aes(total_counts, CumSum)) +
  geom_point() +
  xlab("Filtering Threshold, Minimum Total Counts") +
  ylab("OTUs Filtered") +
  ggtitle("OTUs that would be filtered vs. the minimum count threshold")

pCumSum +
  geom_vline(xintercept = 7000)

library(lattice)
xyplot(CumSum ~ total_counts, data = taxcumsum, 
       xlim = c(0, 900),
       panel = function(x, y,...){
         panel.xyplot(x, y, ...)
         panel.abline(v = 50)
         panel.abline(h = 60)
        }
)

source("taxa_summary.R")

mdt <- fast_melt(ps_dada2)

prevadt <- mdt[, list(Prevalence = sum(count > 0),
                     TotalCounts = sum(count)),
              by = TaxaID]
prevadt %>% 
  ggplot(aes(Prevalence)) +
  geom_histogram() +
  ggtitle("Histogram of Taxa Prevalence")

histogram(~ Prevalence, data = prevadt)
histogram(~ Prevalence, data = prevadt, type = "count")

prevadt[(Prevalence <2), .N]

prevacumsum <- prevadt[, .N, by = Prevalence] %>% 
  setkey(Prevalence) %>% 
  .[, CumSum := cumsum(N)]

## ggplot version
ggplot(prevacumsum, aes(Prevalence, CumSum)) + 
  geom_point()+
  xlab("Filtering Threshold, Prevalence") +
  ylab("OTUs Filtered") +
  ggtitle("OTUs that would be filtered vs. the minimum count threshold")

## lattice version
xyplot(CumSum ~ Prevalence, data = prevacumsum,
       xlab = "Filtering Threshold, Prevalence",
       ylab = "OTUs Filtered", 
       main = "OTUs that would be filtered vs. the minimum count threshold" )

## base r version
plot(CumSum ~ Prevalence, data = prevacumsum,
       xlab = "Filtering Threshold, Prevalence",
       ylab = "OTUs Filtered", 
       main = "OTUs that would be filtered vs. the minimum count threshold" )

# Prevalence vs Total count scatter plot

ggplot(prevadt, aes(Prevalence, TotalCounts)) +
  geom_point(size =4, alpha =0.75) + 
  scale_y_log10()



plot(TotalCounts ~ Prevalence, 
       data = prevadt, log = "y" )


# filter taxa----

keepTaxa <- prevadt[(Prevalence >= 2 & TotalCounts > 3), TaxaID]

ps_filt <- prune_taxa(keepTaxa, ps_dada2)

ps_filt



```


### Filter taxa
```{r chunk8, warning=FALSE, message=FALSE, error=FALSE }
# filter taxa----
source("taxa_summary.R")

mdt <- fast_melt(ps_dada2)

prevadt <- mdt[, list(Prevalence = sum(count > 0),
                     TotalCounts = sum(count)),
              by = TaxaID]
prevadt %>% 
  ggplot(aes(Prevalence)) +
  geom_histogram() +
  ggtitle("Histogram of Taxa Prevalence")



prevadt[(Prevalence <2), .N]

prevacumsum <- prevadt[, .N, by = Prevalence] %>% 
  setkey(Prevalence) %>% 
  .[, CumSum := cumsum(N)]

## ggplot version
ggplot(prevacumsum, aes(Prevalence, CumSum)) + 
  geom_point()+
  xlab("Filtering Threshold, Prevalence") +
  ylab("OTUs Filtered") +
  ggtitle("OTUs that would be filtered vs. the minimum count threshold")



# Prevalence vs Total count scatter plot

ggplot(prevadt, aes(Prevalence, TotalCounts)) +
  geom_point(size =4, alpha =0.75) + 
  scale_y_log10()



plot(TotalCounts ~ Prevalence, 
       data = prevadt, log = "y" )

keepTaxa <- prevadt[(Prevalence >= 2 & TotalCounts > 3), TaxaID]

ps_filt <- prune_taxa(keepTaxa, ps_dada2)

ps_filt

```


### Rarefaction visulization
```{r chunk9, warning=FALSE, message=FALSE, error=FALSE, cache=TRUE}

p_rare <- ggrarecurve(
    obj = ps_dada2,
    indexNames=c("Observe","Chao1","ACE"),
    chunks = 300
 ) + theme(legend.spacing.y=unit(0.02,"cm"),
                legend.text=element_text(size=6))

p_rare


```


### Alpha diversity
```{r chunk10, fig.width=12, warning=FALSE, message=FALSE, error=FALSE }

plot_alpha <- function(ps, factorName) {
  
  alphaobj <- get_alphaindex(ps)
  
  print(head(as_tibble(alphaobj)))
  
  
  p_alpha <- ggbox(
               alphaobj, 
               geom="violin", 
               factorNames = factorName
        ) +
             theme(strip.background = element_rect(colour=NA, fill="grey"))
  
  p_alpha
}

plot_alpha(ps_dada2, "gender") + theme(
  axis.text.x = element_text(angle = 45, hjust = 1)
)


```


### Taxaonomy abundance
```{r chunk11, warning=FALSE, message=FALSE, error=FALSE }

otubar <- ggbartax(obj = ps_filt) +
   xlab(NULL) +
   ylab("Relative abundance(%")

otubar

ggbartax(obj = ps_filt, facetNames = "group", count=TRUE) +
   xlab(NULL) +
   ylab("Relative abundance(%)")


phytax <-  get_taxadf(obj = ps_filt, taxlevels = 2)
phybar <- ggbartax(obj=phytax) +
          labs(
            x = NULL, 
            y = "relative abundance (%)",
            title = "relative abundance (%) of Phylum"
          )
         
phybar


phybar2 <- ggbartax(obj=phytax, facetNames="group", count=TRUE) + 
  labs(
        x = NULL, 
        y = "abundance (counts)",
        title = "Abundance(counts) Phylum"
          )
phybar2

```


### PCA and PCoA analysis
```{r chunk13, warning=FALSE, message=FALSE, error=FALSE }
pcaes <- get_pca(obj = ps_dada2, method="hellinger")

ggordpoint(obj = pcaes,
           biplot = TRUE,
           speciesannot = TRUE,
            
           ellipse=TRUE
          ) + 
           scale_color_manual(values=c("#2874C5", "#EABF00"))

pcoares <- get_pcoa(
     obj = ps_dada2,
     distmethod = "euclidean", 
     method = "hellinger"
)
           
pcoaplot <- ggordpoint(
      obj=pcoares, 
      biplot=TRUE, 
      speciesannot=TRUE,
      #factorNames=c("time"), 
      ellipse=TRUE
      ) +
   scale_color_manual(values=c("#2874C5", "#EABF00"))

pcoaplot

```


### Hierarchial cluster analysis
```{r chunk14, warning=FALSE, message=FALSE, error=FALSE }
hcsample <- get_clust(
  obj = ps_dada2, 
  distmethod = "euclidean",
  method = "hellinger",
  hclustmethod = "average"
)

clustplot1 <- ggclust(obj=hcsample,
                      layout = "rectangular",
                      pointsize=1,
                      fontsize=0,
                      #factorNames=c("time")
                      ) +
              scale_color_manual(values=c("#2874C5", "#EABF00")) +
              theme_tree2(legend.position="right",
                          plot.title = element_text(face="bold", lineheight=25,hjust=0.5))
clustplot1

# circular layout
clustplot2 <- ggclust(obj=hcsample,
                      layout = "circular",
                      pointsize=2,
                      fontsize=4,
                      #factorNames=c("time")
                      ) +
              scale_color_manual(values=c("#2874C5", "#EABF00")) +
              theme(legend.position="right")
clustplot2


```


## Biomarker discovery
```{r chunk15, warning=FALSE, message=FALSE, error=FALSE }
# https://yulab-smu.top/MicrobiotaProcessWorkshop/articles/MicrobiotaProcessWorkshop.html
# 
#This package provides diff_analysis to detect the biomarker. And the results can be visualized by ggdiffbox, ggdiffclade, ggeffectsize and ggdifftaxbar


deres <- diff_analysis(obj = ps_dada2,
                       classgroup = "gender",
                       mlfun = "lda",
                       fitermod = "pvalue",
                       firstcomfun = "kruskal_test",
                       firstalpha = 0.05,
                       strictmod = TRUE,
                       secondcomfun = "vilcox_text",
                       subclmin = 3,
                       subclwilc = TRUE,
                       secondalpha = 0.01,
                       lad =3
                       )


```


```{r chunk16, warning=FALSE, message=FALSE, error=FALSE }

sessionInfo()

```

